---
title: "simpleRegress"
author: "Maximiliano Ortiz"
date: "Fall 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model WEIGHT ~ NECK  this is an example  

```{r}
simp <- lm(WEIGHT~NECK,data=bears)
plot(WEIGHT~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict WEIGHT other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
simp <- lm(WEIGHT~CHEST,data=bears)
plot(WEIGHT~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted overall weight of a bear based on chest has an adjusted R-Squared value of 0.9262, and a Residual standard error value of 33.08. 

```{r}
simp <- lm(WEIGHT~LENGTH,data=bears)
plot(WEIGHT~LENGTH,data=bears)
abline(simp)
summary.lm(simp)
```


The predicted overall weight of a bear based on length has an adjusted R-Squared value of 0.7423, and a Residual standard error value of 61.83.

```{r}
simp <- lm(WEIGHT~HEADWTH,data=bears)
plot(WEIGHT~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted overall weight of a bear based on head width has an adjusted R-Squared value of 6.6064, and a Residual standard error value of 76.41.

```{r}
simp <- lm(WEIGHT~HEADLEN,data=bears)
plot(WEIGHT~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```


The predicted length of a bear compared to head width has an adjusted R squared value of 0.76064, and residual standard error value of 76.41.

```{r}
simp <- lm(WEIGHT~AGE,data=bears)
plot(WEIGHT~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

The predicted overall weight of a bear based on age has an adjusted R-Squared value of 0.5526, and a Residual standard error value of 81.47

##Conclusion and Ranking

According to the data I collected for the adjusted R squared and Residual standard error values when I compared weight to all other numerical values, the ranking is as such. 

1. Weight~Chest: RSE: 33.08 and ARS:0.9262
2. Weight~Neck: RSE: 43.91 and ARS: 0.87
3. Weight~Length RSE: 61.83 and ARS: 0.7423
4. Weight~Head length RSE: 67.81 and ARS: 0.69
5. Weight~Head width RSE: 76.41 and ARS: 0.6064
6. Weight~Age RSE: 81.47 and ARS: 0.5526


